---
title: "PA1_template"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='Figures/')
```

Upzip of the file after the download of the file and Adding the file to the data.frame
```{r}
unzip("activity.zip")
activity1 <- read.csv("activity.csv")
```

Create a histogram of the aggregation for the Histogram of the total number of steps taken each day

```{r}
hist((aggregate(steps ~ date,activity1,sum,na.action = NULL,na.rm = TRUE))$steps,main = "Histogram of total steps taken per day",xlab = "Total steps taken per day")
```

Calculating the total number of steps taken and then there mean and median

```{r}
mean((aggregate(steps ~ date,activity1,sum,na.action = NULL,na.rm = TRUE))$steps)
median((aggregate(steps ~ date,activity1,sum,na.action = NULL,na.rm = TRUE))$steps)
```

Plots to create time series plot with the interval and average number of steps taken.

```{r}
plot.ts(aggregate(steps ~ interval,activity1,mean))
plot(aggregate(steps ~ interval,activity1,mean),type = 'l')
```

Code for the maximum number of steps taken in a particular 5 sec interval. Interval given below

```{r}
aggregate(steps ~ interval,activity1,mean)[which.max(aggregate(steps ~ interval,activity1,mean)$steps),]$interval
```
Code for the Maximum number of steps taken in a particular 5 sec interval, steps and Interval given below.
```{r}
aggregate(steps ~ interval,activity1,mean)[which.max(aggregate(steps ~ interval,activity1,mean)$steps),]
```

Code for calculating the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(rowSums(is.na(activity1)))
```

Strategy to fill or impute the dataset on the missing values with the median.

```{r}
activitiy1AddMissingValue  <- transform(activity1, steps = ifelse(is.na(steps), round(median(steps, na.rm=TRUE)), steps))
sum(rowSums(is.na(activitiy1AddMissingValue)))
```

Histogram after the missing values added with the median.

```{r}
hist((aggregate(steps ~ date,activitiy1AddMissingValue,sum,na.action = NULL,na.rm = TRUE))$steps,main = "Histogram of total steps taken per day with missing value filled in",xlab = "Total steps taken per day")
```

It looks there is no difference in the plot, let's look into the difference in the mean, median before adding the missing values and after filling the missing values.

```{r}
mean((aggregate(steps ~ date,activity1,sum,na.action = NULL,na.rm = TRUE))$steps) - mean((aggregate(steps ~ date,activitiy1AddMissingValue,sum,na.action = NULL,na.rm = TRUE))$steps)
```

```{r}
median((aggregate(steps ~ date,activity1,sum,na.action = NULL,na.rm = TRUE))$steps) - median((aggregate(steps ~ date,activitiy1AddMissingValue,sum,na.action = NULL,na.rm = TRUE))$steps)
```


Adding the Weekdays factor and creating another dataset with the factor
```{r}
activityfactorweekday0rweekend <- factor(weekdays(as.Date(activitiy1AddMissingValue$date)) %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'),levels = c(FALSE,TRUE),labels = c('weekend','weekday'))
summary(activityfactorweekday0rweekend)
```
Adding another dataset with the weekday and weekend column and Adding the necessary packages and also creating the plot for the Weekend and week day information.

```{r}
activitymissingtest <- activitiy1AddMissingValue
activitymissingtest$WeekdayorWeekend <- activityfactorweekday0rweekend
library(dplyr)
activityAggregate <- activitymissingtest  %>% group_by(WeekdayorWeekend, interval) %>% summarize(stepsdetails = mean(steps))
library(ggplot2)
ggplot(activityAggregate, aes(interval,stepsdetails))+geom_line(color="blue")+ facet_wrap(~WeekdayorWeekend, ncol=1)
```
